A boundary-integral framework to simulate \\ the viscous erosion of a porous medium

### 2018-04-06 docs/methods/erosion_methods.tex
NM

Flow-induced erosion deteriorates and reshapes solid material over a range of scales found in nature, from large-scale land formations \cite{han1969, Rothman2012, coh-dev-sey-yi-szy-rot2015} down to human-scale features and patterns \cite{daerr2003erosion}. Though less visible, these same forces are at work at the very smallest scales, slowly deteriorating the individual constituents of porous media (e.g.~solid, sand, or clay) or biological structures like plaque and biofilms \cite{pic-van-hei2000, sha2002, gro-gij-van-fer-hat-van-yua-wen2007}. Motivated by such small-scale applications, we develop here numerical methods to simulate fluid-mechanical erosion of multiple solid bodies in Stokes flow --- the most relevant regime for groundwater and bio-fluid applications.

Recent work in the high-Reynolds regime has highlighted the importance of the feedback between flow and morphology that occurs during erosion \cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013, hewett2017evolution, moore2017riemann}. A similar feedback mechanism is at work in the related processes of dissolution and melting \cite{Huang2015, kondratiuk2015steadily, rycroft2016asymmetric, cohen2016erosion, hewett2017pear, claudin2017dissolution}. In such cases, the flow alters the shape of the immersed structures, which in turn modifies the surrounding flow, and so on. Collectively, such processes have been referred to as `sculpting with flow'~\cite{ristroph2018sculpting}. From this viewpoint, the example of an eroding porous medium presents a new opportunities and challenges. First, the medium itself is composed of many solid constituents, and so multi-body interactions, especially close interactions, must be considered. Second, the long-range interactions facilitated by Stokes flow even more strongly couples the shape evolution of multiple bodies.

The methods developed here will allow us to simulate the erosion of many bodies, as demonstrated in Fig.~\ref{fig:50bodies} in for a medium composed of 50 bodies with randomly assigned initial positions and sizes. In this work, mechanical erosion arises from the shear-stress induced by the flow, as has been experimentally confirmed \cite{ris-moo-chi-she-zha2012}. In Fig.~\ref{fig:50bodies}, color represents the vorticity of the surrounding flow. When evaluated on the solid boundaries, vorticity is equivalent to shear, and thus provides a convenient tool for visualizing local erosion rates. As shown in the figure, our simulations allows solid bodies to diminish in size and vanish in finite time, while other larger bodies persist.

@@

\vsp{2}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Contributions} The present study develops numerical methods
to simulate erosion of multiply-connected two-dimensional geometries.
In this fashion, individual pores are resolved rather than
homogenized.  A boundary integral equation (BIE) is used to solve the
incompressible Stokes equations in the complex geometry.  The resulting
method resolves individual grains and the non-negligible interactions
between bodies.  To link the eroding body and fluid dynamics, the body
is eroded in the normal direction at a rate that is proportional to the
magnitude of the shear stress.

>>>>

Flow-induced erosion deteriorates and reshapes solid material over a range of scales found in nature, from large geological formations \cite{han1969, Rothman2012, coh-dev-sey-yi-szy-rot2015}, down to human-scale features and patterns \cite{daerr2003erosion}. Though less visible, these same forces are at work at the very smallest scales, slowly deteriorating the individual constituents of porous media (e.g.~solid, sand, or clay) or biological structures like plaque and biofilms \cite{pic-van-hei2000, sha2002, gro-gij-van-fer-hat-van-yua-wen2007}. Motivated by such applications, we develop here numerical methods to simulate fluid-mechanical erosion of multiple solid bodies in Stokes flow --- the most relevant regime for groundwater and bio-fluid applications.

Recent work in the high-Reynolds regime has highlighted the importance of the feedback between flow and morphology that occurs during erosion \cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013, hewett2017evolution, moore2017riemann}, and a similar feedback occurs in the related processes of dissolution and melting \cite{Huang2015, kondratiuk2015steadily, rycroft2016asymmetric, cohen2016erosion, hewett2017pear, claudin2017dissolution}. In such cases, the flow alters the shape of the immersed structures, which in turn modifies the surrounding flow, and so on. Collectively, such processes have been referred to as `sculpting with flow'~\cite{ristroph2018sculpting}. From this viewpoint, the example of an eroding porous medium presents a new opportunities and challenges. First, the medium itself is composed of many solid constituents, and so multi-body interactions, especially close interactions, must be considered. Second, the long-range interactions facilitated by Stokes flow even more strongly couples the shape evolution of neighboring bodies, making their interactions even more important.

\paragraph{Contributions}
The methods developed here will allow us to simulate the erosion of many bodies, as demonstrated in Fig.~\ref{fig:50bodies} for the case of 50 bodies (of randomly assigned initial sizes and positions). In this work, erosion of solid material arises from the mechanical stress induced by the surrounding flow. In particular, we take the local rate of erosion to be proportional to the magnitude of the shear stress -- a law that has been verified by laboratory experiments \cite{ris-moo-chi-she-zha2012}. As seen in the figure, erosion not only diminishes the size of the bodies, but also alters their shapes considerably. Notice, in particular, that many bodies develop corner-like features and nearly flat faces, with straight channels intervening between them. In the figure, color represents the vorticity of the surrounding flow, which, evaluated on solid boundaries, reduces to shear. In this way, vorticity provides a convenient tool for visualizing local erosion rates.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The present study develops numerical methods
to simulate erosion of multiply-connected two-dimensional geometries.
In this fashion, individual pores are resolved rather than
homogenized.  A boundary integral equation (BIE) is used to solve the
incompressible Stokes equations in the complex geometry.  The resulting
method resolves individual grains and the non-negligible interactions
between bodies.  To link the eroding body and fluid dynamics, the body
is eroded in the normal direction at a rate that is proportional to the
magnitude of the shear stress.
